---
title: "Car Accidents by Car Type"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
    theme: yeti
---

```{r setup, include=FALSE, echo = FALSE}
library(tidyverse)
library(lubridate)
```

```{r, include=FALSE, message=FALSE, warning=FALSE}
source("code/data_cleaning.R")
source("code/general_formatting.R")
```

### Car Accident by Car Type

Then, we want to know how types of cars are associated with the number of accidents. A bar plot of the number of accidents by car type clearly represents the relationship.  

### Number of Accidents by Car Type

```{r, message=FALSE, warning=FALSE}
accidents2 = accidents1 %>% 
  group_by(vehicle_type_code_1) %>% 
  summarize(obs = n()) %>% 
  arrange(desc(obs))

accidents_cartype_bar = 
  accidents2 %>% 
  filter(obs > 100) %>%
  filter(!is.na(vehicle_type_code_1)) %>% 
  filter(obs > 500) %>% 
  mutate(vehicle_type = fct_reorder(vehicle_type_code_1, obs)) %>% 
  ggplot(aes(x = vehicle_type, y = obs, fill = vehicle_type)) +
  geom_bar(stat = "identity") + 
  theme_classic() + 
  labs(title = "Number of Accidents by Car Type", 
       y = "Number of Accidents", 
       x = "Car Type") +
  geom_text(size = 3, aes(label = obs), position = position_dodge(width=0.9), vjust=-0.25) + 
  ylim(0, 45000)

accidents_cartype_bar + theme(axis.text.x=element_blank(), axis.ticks.x=element_blank())
```

The plot shows the frequency (>500) of accidents by car type. There are mainly two types of cars with much higher frequencies than other types: sedan and station wagon/sport utility vehicle. The frequencies of these two types in 2020 are more than 25000. Other common types are taxi, pick-up truck, and box truck. The frequencies of these three types are more than 1000. Other types are bus, bike, tractor truck diesel, and motorcycles also have a frequency higher than 500.
