---
title: "Mapping"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
    theme: yeti
---
```{r setup, include=FALSE}
library(tidyverse)
library(leaflet)
library(plotly)
library(dplyr)
library(lubridate)
knitr::opts_chunk$set(
  fig.width = 10,
  fig.asp = .5,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

```{r, include=FALSE, message=FALSE, warning=FALSE}
source("code/data_cleaning.R")
source("code/general_formatting.R")
```

## Mapping: Number of car accidents
To more clearly visualize the relationship and between latitude and longitude and the number of crashes, we first aggregated the latitude and longitude and kept two decimal places. 

From the mapping plot, we can see that from 2020/1 to 2020/8, the number of accidents occurred more in Bronx and Manhattan, followed by Brooklyn. Fewer accidents occurred in Queens and Staten Island. 
```{r, message = FALSE, warning=FALSE}
NYC_data1 = read_csv("./data/NYC Accidents 2020.csv",
                col_names = TRUE) %>%
  janitor::clean_names() 

map_data =
NYC_data1 %>%
  filter(!is.na(borough)) %>%
  filter(!is.na(latitude)) %>%
  filter(!is.na(longitude)) %>%
  mutate(
    latitude = round(latitude, digits = 2),
    longitude = round(longitude, digits = 2)) %>%
  group_by(latitude, longitude) %>%
  count(latitude, longitude, borough) %>%
  arrange(desc(n)) %>%
mutate(text_label = str_c("</b>Lat: ", latitude, "°", "</b><br>Lng: ", longitude, "°", "</b><br>Borough: ", borough, "</b><br>Number_of_accidents: ", n))

pal = colorNumeric(
  palette = "YlGnBu",
  domain = map_data$n) 

map_data %>%
  leaflet()  %>%
  setView(lng = -74.00666, lat = 40.71643, zoom = 10) %>% 
  addTiles() %>% 
  addLegend("bottomright", 
            pal = pal, 
            values = ~n,
            title = "Number of car accidents</br>",
            bins = 10,
            opacity = 1,
            labFormat = labelFormat(suffix = " cases")
  ) %>%
  addCircleMarkers(
    lng = ~longitude,
    lat = ~latitude,
    color = ~pal(n),
    radius = 4,
    popup = ~ text_label,
    fillOpacity = 0.8,
    stroke = FALSE,
    opacity = 1) 
```


## Mapping: Number of Injuries
To visualize the number of injuries in each accidents, we employ the latitude and longitude of each accidents to see where have the most injuries.  

From the plot, we can tell that the maximum injuries in one accident is 30 people. But most accidents have injuries less than 5. A few accidents have more than 5 injuries.  
```{r, message = FALSE, warning=FALSE}
accidents_2020_map = NYC_data1 %>% 
  filter(!is.na(borough)) %>% 
  filter(!is.na(latitude)) %>%
  filter(!is.na(longitude)) %>%
  mutate(
    total_injured = number_of_persons_injured + number_of_pedestrians_injured + number_of_cyclist_injured + number_of_motorist_injured,
    textlab = paste0("Number of Injuries: ", total_injured, "\nAddress: ", on_street_name)
  )
```

```{r, message = FALSE, warning=FALSE}
total_injuried_map = accidents_2020_map %>% 
  plot_ly(
    lat = ~latitude, 
    lon = ~longitude, 
    type = "scattermapbox", 
    mode = "markers", 
    alpha = 0.5,
    color = ~total_injured,
    text = ~textlab,
    colors = "YlOrRd")
```


```{r, message = FALSE, warning=FALSE}
total_injuried_map %>% layout(
    mapbox = list(
      style = 'carto-positron',
      zoom = 9,
      center = list(lon = -73.9, lat = 40.7)),
      title = "<b> Total Number of Injuries </b>",
      legend = list(title = list(text = "Borough of Accidents", size = 9),
                    orientation = "h",
                   font = list(size = 9)))
```
