---
title: "Statistical Analyses"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
    theme: yeti
---
```{r setup, include=FALSE}
library(tidyverse)
library(dplyr)
library(lubridate)
library(rvest)
library(knitr)
```

```{r, include=FALSE, message=FALSE, warning=FALSE}
source("code/data_cleaning.R")
source("code/general_formatting.R")
```

## Chi-Squared test - Car accidents in different weekdays and different boroughs
To test whether the proportion of numbers of car accidents in each weekday among boroughs are equal, we perform Chi-Squared test.

***H0***: The proportion of car accidents number in weekdays among boroughs are equal.    
***H1***: Not all proportion of car accidents number in weekdays among boroughs are equal.
```{r, message = FALSE, warning=FALSE}
NYC_data1 = read_csv("./data/NYC Accidents 2020.csv",
                col_names = TRUE) %>%
  janitor::clean_names() 

week_accidents = 
  NYC_data1 %>%
  dplyr::select(crash_date, borough) %>%
  mutate(weekdays = weekdays(NYC_data1$crash_date, abbreviate = T) ,
         weekdays = fct_recode(weekdays, "Mon" = "周一", "Tue" = "周二", "Wed" = "周三", "Thu" = "周四", "Fri" = "周五", "Sat" = "周六", "Sun" = "周日")) %>%
  filter(!is.na(borough)) %>%
  mutate(weekdays = as.factor(weekdays),
         weekdays = fct_relevel(weekdays, "Mon", "Tue", "Wed", "Thu", "Fri", "Sat", "Sun"))

table(week_accidents$borough, week_accidents$weekdays)

chisq.test(table(week_accidents$borough, week_accidents$weekdays))

x_crit = qchisq(0.95, 24)
x_crit
```

**Interpretation**: At significant level α=0.05, p-value = 6.303e-07 < 0.05, so we reject the null hypothesis and conclude that there is at least one borough’s proportion of car accidents for weekdays is different from others.


## Chi-square test - Each borough Car type’s proportion of accident amounts

To test whether the proportion of numbers of car accidents in each five car types among boroughs are equal, we performed Chi-square test.   

***H0***: Each borough’s proportion of accident amounts for five car types are same.   
***H1***: Each borough’s proportion of accident amounts for five car types are not same.   
```{r, message = FALSE, warning=FALSE}
accidents_2020 = read_csv("./data/NYC Accidents 2020.csv") %>% 
  janitor::clean_names()

five_common_cartype = 
  accidents_2020 %>%
  select(borough, vehicle_type_code_1) %>% 
  filter(vehicle_type_code_1 %in%
           c("Sedan",
             "Station Wagon/Sport Utility Vehicle",
             "Taxi",
             "Pick-up Truck",
             "Box Truck")) %>%
  count(vehicle_type_code_1, borough) %>% 
  pivot_wider(
    names_from = "vehicle_type_code_1",
    values_from = "n"
  )  %>% 
  data.matrix() %>% 
  subset(select = -c(borough))

rownames(five_common_cartype) <- c("Bronx", "Brooklyn", "Manhattan", "Queens", "Staten Island", "Others")

five_common_cartype %>% 
  knitr::kable(caption = "Table of Top Five Car Type", caption.pos = "top")
```

```{r, message = FALSE, warning=FALSE}
chisq.test(five_common_cartype)
```
**Interpretation**: The result of chi-square shows that χ2>χcrit, at significant level α=0.05, so we reject the null hypothesis and conclude that there does exist at least one car type's proportion of accident amounts is different than others.



## Proportion test - The proportions of car accidents across boroughs. 

We want to see whether the car accidents rates are same of each borough, so we conduct a proportion test. We derived the population of each borough from the most recent census.  

***H0***: The proportion of the car accidents is the same across all boroughs.  
***H1***: The proportion of the car accidents is not the same across all boroughs.  
```{r, message = FALSE, warning=FALSE}
data <- read.csv("./data/NYC Accidents 2020.csv") %>% 
  janitor::clean_names() %>% 
  select(crash_date, crash_time, borough, number_of_persons_injured, 
  number_of_persons_killed, number_of_pedestrians_injured, 
  number_of_pedestrians_killed, number_of_cyclist_injured, 
  number_of_cyclist_killed, number_of_motorist_injured, 
  number_of_motorist_killed) %>% 
#delete the NAs in borough
  filter(borough %in% c("BRONX", "BROOKLYN", "QUEENS", "MANHATTAN", "STATEN ISLAND"))

url = "https://www.citypopulation.de/en/usa/newyorkcity/"
nyc_population_html = read_html(url)

population = nyc_population_html %>% 
  html_elements(".rname .prio2") %>% 
  html_text()

boro = nyc_population_html %>% 
  html_elements(".rname a span") %>% 
  html_text()

nyc_population = tibble(
  borough = boro,
  population = population %>% str_remove_all(",") %>% as.numeric()
) 
  
car_accident = data %>%
  count(borough) %>% 
  mutate(borough = str_to_title(borough))

acci_popu_boro = left_join(car_accident, nyc_population)

acci_popu_boro %>% 
  knitr::kable(caption = "Results Table", caption.pos = "top")

prop.test(acci_popu_boro$n, acci_popu_boro$population)
```
**Interpretation**: From the test result, we can see that the p-value is smaller than 0.01, so we have enough evidence to conclude that the proportions of car accidents are different across boroughs.



## ANOVA Test - Month and accidents

In order to study how month are associated with the number of car accidents, We try to use an ANOVA test across months.

***H0***: The average number of accidents are not different across months.    
***H1***: The average number of accidents are different across months.    
```{r, message = FALSE, warning=FALSE}
accidents = 
  read_csv("./data/NYC Accidents 2020.csv") %>% 
  janitor::clean_names() %>% 
  mutate(
    borough = case_when(borough == "BRONX" ~ "Bronx", 
                        borough == "BROOKLYN" ~ "Brooklyn", 
                        borough == "QUEENS" ~ "Queens", 
                        borough == "MANHATTAN" ~ "Manhattan", 
                        borough == "STATEN ISLAND" ~ "Staten Island"), 
    crash_date = as.Date(crash_date, format = "%y-%m-%d"), 
    weekday = weekdays(crash_date), 
    month = month(crash_date), 
    day = day(crash_date)
  ) 

fit_accidents = 
  accidents %>% 
  mutate(month = as.factor(month)) %>% 
  group_by(month, weekday, day) %>% 
  dplyr::summarize(num_accidents = n()) 
fit_accidents_month = lm(num_accidents ~ month, data = fit_accidents)  
anova(fit_accidents_month) %>% 
  knitr::kable(caption = "One way anova of number of accidents and month", caption.pos = "top")
```

**Interpretation**: As indicated by the result of the ANOVA test, the p value is very small. Therefore, the null hypothesis is rejected and we can conclude that the average number of accidents are different across months in New York City in 2020.
