---
title: "Statistical Analyses"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
    theme: yeti
---
```{r setup, include=FALSE}
library(tidyverse)
library(dplyr)
library(lubridate)
library(rvest)
library(knitr)
```

```{r, include=FALSE, message=FALSE, warning=FALSE}
source("code/data_cleaning.R")
source("code/general_formatting.R")
```

### Chi-Squared test - Car accidents on weekdays among boroughs

To test whether the proportions of car accidents in each weekday among boroughs are equal, we perform the Chi-Squared test.

***H0***: The proportions of car accidents on weekdays among boroughs are equal.

***H1***: Not all proportions of car accidents on weekdays among boroughs are equal.

```{r, message = FALSE, warning=FALSE}
week_accidents = 
  accidents1 %>%
  dplyr::select(crash_date, borough) %>%
  mutate(weekdays = weekdays(accidents1$crash_date, abbreviate = T)) %>% 
  filter(!is.na(borough)) %>%
  mutate(weekdays = as.factor(weekdays),
         weekdays = fct_relevel(weekdays, "Mon", "Tue", "Wed", "Thu", "Fri", "Sat", "Sun"))

table(week_accidents$borough, week_accidents$weekdays)

chisq.test(table(week_accidents$borough, week_accidents$weekdays))

x_crit = qchisq(0.95, 24)
x_crit
```

**Interpretation**: At significant level $\alpha$ = 0.05, $p-value$ = 6.303e-07 < 0.05, so we reject the null hypothesis and conclude that there is at least one borough’s proportion of car accidents for weekdays different from others.

### Chi-square test - Car type’s proportion of accident amounts among boroughs

To test whether the proportions of car accidents in five car types among boroughs are equal, we performed the Chi-square test.   

***H0***: Proportions of accident amounts for five car types are equal among boroughs.   
***H1***: Not all proportions of accident amounts for five car types are not equal among boroughs.   

```{r, message = FALSE, warning=FALSE}
five_common_cartype = 
  accidents1 %>%
  select(borough, vehicle_type_code_1) %>% 
  filter(vehicle_type_code_1 %in%
           c("Sedan",
             "Station Wagon/Sport Utility Vehicle",
             "Taxi",
             "Pick-up Truck",
             "Box Truck")) %>%
  count(vehicle_type_code_1, borough) %>% 
  pivot_wider(
    names_from = "vehicle_type_code_1",
    values_from = "n"
  )  %>% 
  data.matrix() %>% 
  subset(select = -c(borough))

rownames(five_common_cartype) <- c("Bronx", "Brooklyn", "Manhattan", "Queens", "Staten Island", "Others")

five_common_cartype %>% 
  knitr::kable(caption = "Table of Top Five Car Type", caption.pos = "top")
```

```{r, message = FALSE, warning=FALSE}
chisq.test(five_common_cartype)
```

**Interpretation**: At significant level $\alpha$ = 0.05, the result of chi-square shows that $\chi^2$ > $\chi_{crit}$, so we reject the null hypothesis and conclude that there is at least one car type's proportion of accident amounts different from others.

### Proportion test - The proportions of car accidents among boroughs

We want to see whether the car accident rates are the same among boroughs, so we conduct a proportion test. We obtained the population of each borough from the most recent census.  

***H0***: The proportions of the car accidents are equal among boroughs.  

***H1***: The proportions of the car accidents are not equal among boroughs.  

```{r, message = FALSE, warning=FALSE}
url = "https://www.citypopulation.de/en/usa/newyorkcity/"
nyc_population_html = read_html(url)

population = nyc_population_html %>% 
  html_elements(".rname .prio2") %>% 
  html_text()

boro = nyc_population_html %>% 
  html_elements(".rname a span") %>% 
  html_text()

nyc_population = tibble(
  borough = boro,
  population = population %>% str_remove_all(",") %>% as.numeric()
) 
  
car_accident = accidents1 %>%
  filter(!is.na(borough)) %>%
  count(borough) %>% 
  mutate(borough = str_to_title(borough))

acci_popu_boro = left_join(car_accident, nyc_population)

acci_popu_boro %>% 
  knitr::kable(caption = "Results Table", caption.pos = "top")

prop.test(acci_popu_boro$n, acci_popu_boro$population)
```

**Interpretation**: From the test result, we can see that the $p-value$ is smaller than 0.01, so we have enough evidence to conclude that the proportions of car accidents are different across boroughs.

### ANOVA Test - Month and accidents

In order to study how month is associated with the number of car accidents, we try to use an ANOVA test across months.

***H0***: The average numbers of accidents are equal across months.   

***H1***: The average numbers of accidents are not equal across months.   

```{r, message = FALSE, warning=FALSE}
fit_accidents = 
  accidents1 %>% 
  mutate(month = as.factor(month)) %>% 
  group_by(month, weekday, day) %>% 
  dplyr::summarize(num_accidents = n()) 
fit_accidents_month = lm(num_accidents ~ month, data = fit_accidents)  
anova(fit_accidents_month) %>% 
  knitr::kable(caption = "One way anova of number of accidents and month", caption.pos = "top")
```

**Interpretation**: As indicated by the result of the ANOVA test, the $p-value$ is very small. Therefore, the null hypothesis is rejected and we can conclude that the average numbers of accidents are different across months in New York City in 2020.
